---
title: "Lie/numbers analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Decsciption and aim of this markdown

The aim of this markdown is replicate Valter's analysis of SNARC data using the R programming language. R provides a flexible way to analyse and visualise data. The markdown format allows for the analyse code to be separted into specific sections, and saved in html format. Before I continue into the analysis, I will remind both myself and Valter of the experiment:

## Experiment description - the context

Participants are told that they are hackers that have stolen codes. The police want to crack the code but participants have to communicate codes to their lawyers without being caught.

## Control_SNARC 

Participants shown just picture of the lawyer and a number. Participants have to respond with equivalent number (odd or even). This is the standard SNARC with context of being in a court.

## Lie_SNARC

In the lie condition, sometimes the picture is the lawyer and other times it is the cop. When it is the cop, they have to lie (i.e. repond with the opposite key, if the number is odd they had to respond as if it is  even). When the picture is a lawyer, they have to tell the truth (i.e. respond with the correct key). Here 2/3 trials are lawyer (standard SNARC) whereas the 1/3 of trials participants had to lie (does this have any effect on the SNARC?)

## What is being investigated?

SNARC effect - Spatial-Numerical Association of Codes effect proposes faster reaction times to small/large numbers on the left/right hand side.

MARC effect - Markedness of Response Codes effect proposes faster reaciton times for left/right hand side repsonses to odd/even numbers

Odd effect - faster reaction times to even numbers, slower reaction times for odd numbers. 

## Understanding column names and data types from the raw data

Key: a = left key, g = right key 

Accuracy: 1 = correct, 0 = incorrect

## A brief description of the packages I will be using

To manipulate dataframes for analysis and data visualisations, I will be using a package called *dplyr*. This package provides functions such as *dplyr::filter()* (to filter by specific values), *dplyr::group_by()* (to group by a specific variable before performing a calculation) and others. This package also has something called a pipe (%>%). This symbol essential means "and then do this" which allows you to manipulate the dataframe before using it.

It works in concordance with other packages (ggplot2) so you can manipulate dataframes to visualise data, without having to create lots of new dataframes from scratch. 

## Loading packages

```{r echo = TRUE,  message = 'hide', warning = 'hide', results = 'hide'}
# rm(list = ls())
library(ggplot2) # for plotting
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(readxl) # reading excel files 
library(openxlsx) # loads data sets from multiple sheets
library(car) # inferential stats
library(MASS) # inferential stats
library(EnvStats) # inferential stats
library(Rmisc) # confidence intervals 
library(sjstats) # for calculating effect sizes for ANOVA
library(effsize) # cohen's D function
library(lme4) # for linear mixed effects modelling
library(lmtest) # testing linear regression models
library(fitdistrplus)
```

Above I am loading the packages that will be necessary for the analysis. Each package has specific utilities that will be needed to perform analysis. It is important to load them at the start of your script otherwise you will encounter errors.

## Loading data

```{r echo = TRUE, message = 'hide', warning = 'hide', results = 'hide'} 
# dput(head(df,n)) provides example of data structure for stack overflow

setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/valter/")

dat <- loadWorkbook("Data_Lie_Numbers_Clean.xlsx")
sheetNames <- sheets(dat)
for(i in 1:length(sheetNames)){
  assign(sheetNames[i], readWorkbook(dat, sheet = i))
}
```

Here I am loading the data. I set my working directory using the *setwd()* function, and then load in data from both sheets of the excel file.

## Checking for na values

```{r}
Lie_SNARC %>%
  dplyr::select(numbers, accuracy, RTs) %>%
  dplyr::filter(is.na(RTs))

Control_SNARC %>%
  dplyr::select(numbers, accuracy, RTs) %>%
  dplyr::filter(is.na(RTs))
```

The first thing I do when I get a new dataset is check for na values. If these are not sorted, they can become an issue when analysing the data. I use the *is.na()* function to tell me where and how many na values each dataset has. In the *Lie_SNARC* dataset, I have 43 trials with an na value within the RT column. For the *Control_SNARC* dataset, I have 1 trial with an value within the RT column.

To further investigate, I filter both dataframes to look at the trials that have na values. For each na, the accurary is 0. I assume what has happened is the subject has ran out of time to respond or inputted an invalid resposne thus they do not get an RT and the accuracy is indicated as an error.

**Dealing with NA values**

When focusing on just error analysis, I shall keep the na values as they provide interesting information (i.e. conditions where subjects did no provide a response). For the MARC and SNARC analysis, I will have to remove the na values as there is no *response_hand* if no response is recorded.

## Renaming conditions

```{r}
# cleaning control condition
Control_SNARC <- Control_SNARC %>%
  dplyr::select(numbers, condition, key, accuracy, RTs, participant) %>%
  arrange(participant) %>%
  mutate(condition = "controltrue")

# cleaning experimental conditon
Lie_SNARC <- Lie_SNARC %>%
  dplyr::select(numbers, condition, key, accuracy, RTs, participant) %>%
  arrange(participant) %>%
  dplyr::mutate(condition = case_when(condition == "t" ~ "exptrue",
                                      condition == "l" ~ "explie"))
workingdata <- rbind(Control_SNARC, Lie_SNARC)
```

The piece of code above is mainly for my own benefit - I am just renaming conditions. From the *Control_SNARC* dataframe, I am selecting the important columns (numbers, condition, accuracy, RTs and participant), I then arrange them by participant, and then rename name the condition *"controltrue"*. This indicates it was a control condition and subjects had to tell the truth i.e. the basic SNARC effect.

The *Lie_SNARC* dataframe contains trials from the experimental block - either where subjects still had to tell the truth, or if they had to lie. I found the *t* and *l* names confusing, so I rename them exptrue and explie.

I then combine both dataframes into one overall dataframe called *workingdata*.

## Creating handedness column

```{r}
workingdata <- workingdata %>%
  dplyr::mutate(response_hand = case_when(key == "a" ~ "left",
                                          key == "g" ~ "right"))
```

Again, this is for my own benefit. I am creating a column to replace "key" as I found it confusing. The new column is called hand and denotes the left or right hand to make it easier to understand the dataframe.

The *case_when* function allows me to create a new column of data dependiing on the value of another variable - in this case, *key*.

## Creating parity column

```{r}
workingdata <- workingdata %>%
  dplyr::mutate(parity = ifelse(numbers %% 2 == 0, "even", "odd"))
```

Finally, I create a *parity* column. The *ifelse()* function evaluates the first argument. When the value in the numbers column is divided by 2, if it equals 0 create a new column and insert *even*. For an other value after dividing, it much be odd thus insert *odd*.

I have cleaned the data and set it up in a way I feel comfortable. The data manipulation so far is personal preference, but I will be using the variables created above for the rest of the analysis.

## Overview conditions - (replication of slide 2)

The following code computes average errors and meanRTs for each condition.

```{r}
workingdata$condition <- factor(workingdata$condition, levels = c("controltrue", "exptrue", "explie"))

"error pecentage" 
ggplot(workingdata %>%
  dplyr::group_by(condition, accuracy) %>%
  dplyr::summarise(acc_num = n()) %>%
  dplyr::ungroup() %>%
  dplyr::group_by(condition) %>%
  dplyr::mutate(perc = (acc_num / sum(acc_num)) * 100) %>%
  dplyr::filter(accuracy == 0), aes(x = condition, y = perc)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(perc, 2),  vjust = -1)) +
  ylim(0, 15) +
  ylab("Percentage of errors (%)") +
  scale_x_discrete("Condition", labels = c("controltrue" = "Control", "explie" = "Police", "exptrue" = "Lawyer")) +
  ggtitle("Errors %") +
  theme_gray() 

"errors per condition"
ggplot(workingdata %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::group_by(condition) %>%
         dplyr::summarise(errors = n()), aes(x = condition, y = errors)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = errors, vjust = -1)) +
  ylim(0, 1000) +
  ylab("Numbers of errors") +
  scale_x_discrete("Condition", labels = c("controltrue" = "Control", "explie" = "Police", "exptrue" = "Lawyer")) +
    ggtitle("Errors") +
  theme_gray() 

"mean RTS per  condition"
ggplot(workingdata %>%
         drop_na() %>%
         dplyr::filter(accuracy == 1) %>%
         dplyr::group_by(condition) %>%
         dplyr::summarise(meanRT = mean(RTs)), aes(x = condition, y = meanRT)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(meanRT, 2), vjust = -1)) +
  ylim(0, 1.2) +
  ylab("Mean reaction times (s)") +
  scale_x_discrete("Condition", labels = c("controltrue" = "Control", "explie" = "Police", "exptrue" = "Lawyer")) +
  ggtitle("Mean RT") +
  theme_gray() 
```

The above code computes the graphs for slide 2 in the presentation that you sent to me (and that I sent back to you). The graphs visualise:

- Number of errors per condition (more errors in the condition where the police officer was present)
- Percentage of errors per condition (higher percentage of errors where the police officer was present)
- Mean reaction time per conditon (faster RTs for the control condition i.e. no pictures, just standard SNARC)

## Control condition RTs (lawyer only) - (replication of slide 3)

```{r}
"RTs over different numbers"
ggplot(workingdata %>%
         drop_na() %>%
         dplyr::filter(condition == "controltrue") %>%
         dplyr::group_by(numbers, response_hand) %>%
         dplyr::summarise(meanRT = mean(RTs)), aes(x = numbers, y = meanRT, col = response_hand)) +
  geom_point() +
  geom_line() +
  scale_x_continuous("Numbers", breaks = seq(0, 9, 1)) +
  scale_y_continuous("Mean RT (s)", breaks = seq(0, 0.62, 0.02)) +
  ggtitle("RTs for left and right hands across numbers")

"Odd effect (95% CIs and corresponding t-test"
workingdata$parity <- factor(workingdata$parity, levels = c("odd", "even"))

ggplot(workingdata %>%
         drop_na() %>%
         dplyr::filter(condition == "controltrue") %>%
         summarySE(measurevar = "RTs", groupvars = c("parity")), aes(x = parity, y = RTs)) + 
  geom_col() +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .2) +
  coord_cartesian(ylim = c(0.55, 0.60)) +
  ggtitle("Odd effect RT analysis")

control_ttest_df <- workingdata %>%
  drop_na() %>%
  dplyr::filter(condition == "controltrue") %>%
  dplyr::select(RTs, parity, participant)

control_ttest_df <- reshape(control_ttest_df, idvar = "participant", timevar = "parity", direction = "wide")

t.test(control_ttest_df$RTs.even, control_ttest_df$RTs.odd, paired = TRUE)
cohen.d(control_ttest_df$RTs.even, control_ttest_df$RTs.odd, paired = TRUE)

"MARC effect (95% CIs and corresponding 2x2 ANOVA"
workingdata$parity <- factor(workingdata$parity, levels = c("odd", "even"))
workingdata$response_hand <- factor(workingdata$response_hand, levels = c("left", "right"))

ggplot(workingdata %>%
         drop_na() %>%
         dplyr::filter(condition == "controltrue") %>%
         summarySE(measurevar = "RTs", groupvars = c("parity", "response_hand")), aes(x = parity, color = response_hand, group = response_hand, y = RTs)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  ggtitle("MARC effect RT analysis")

result_MARC <- ezANOVA(workingdata %>%
                         drop_na() %>%
                         dplyr::filter(condition == "controltrue"), dv = RTs, wid = participant, within = .(parity, response_hand), type = 3, detailed = TRUE, return_aov = TRUE)
result_MARC

eta_sq(result_MARC$aov, partial = TRUE)
```

This code chunk aims to replicate slide 3 of the summary analysis.

**Left-Right line graph**

Line graph represents the mean reaction times for each hand responded with (left and right) for various numbers in the control condition. As we might expect, reaction times are faster for congruent hand-parity pairs (left-odd, right-even).

**Odd effect plot and t-test**

The 95% CIs match the example. Mean RTs are slightly different by looking at the degrees of freedom, the number of participants are not the same (perhaps I have a more updated dataset?). Either way, the pattern of results is the same.

The corresponding t-test comparing the 2 means comes out as non-signifcant [t(37) = 0.23, p = 0.81].

**MARC effect plot and 2X2 ANOVA**

Appears to the show a fairly strong parity x hand interaction - RTs are faster when the hand repsonded with and the pairty are conrguent (left-odd, right-even).

2x2 ANOVA reveals significant interaction with large effect size [f(1, 37) = 4.61, p < 0.001, eta^2 = 0.11].

**Consideration - ANOVA versus GLM** 

A standard ANOVA assumes a normal distribution, however RTs rarely follow a normal distribution. GLMs often provide more statistical power when analysing RT data because you can specify an error distribution that fits your data better (such as Gamma or Inverse-Gaussian). This paper details differences in analysing RT data between linear and generalised linear *mixed effects* model however I think the same applies when not specifying mixed effects (https://www.frontiersin.org/articles/10.3389/fpsyg.2015.01171/full)

## Control condition RTs (lawyer only) - (replication of slide 4)

```{r}
"creating magnitude variable"
workingdata <- workingdata %>%
  dplyr::mutate(magnitude = case_when(numbers <= 4 ~ "small",
                                      numbers > 4 ~ "large"))
"SNARC effect  (95% CIs and 2x2 ANOVA)"
workingdata$magnitude <- factor(workingdata$magnitude, levels = c("small", "large"))
workingdata$response_hand <- factor(workingdata$response_hand, levels = c("left", "right"))

ggplot(workingdata %>%
         drop_na() %>%
         dplyr::filter(condition == "controltrue") %>%
         summarySE(measurevar = "RTs", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = RTs)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  ggtitle("SNARC effect RT analysis")

result_SNARC <- ezANOVA(workingdata %>%
                          drop_na() %>%
                          dplyr::filter(condition == "controltrue"), dv = RTs, wid = participant, within = .(magnitude, response_hand), type = 3, detailed = TRUE, return_aov = TRUE)
result_SNARC

eta_sq(result_SNARC$aov, partial = TRUE)
```

Firstly I make dataframes *small* and *large* in order to compute a new variable called magnitude. This will be used (alongside hand) to analyse the SNARC effect

**2x2 factorial ANOVA**

Significant main effect of magnitude (smaller number reacted to faster) however more importantly, a significant interaction with large effect size [f(1, 37) = 4.64, p < 0.001, eta^2 = 0.11].

This indicates a classic SNARC in the only lawyer control condition (i.e. when responding to larger numbers with right hard, RTs are faster. When responding to small numbers with left hand, RTs are faster). 

##  Control condition errors (lawyer only) - Odd effect error analysis (assessing distribution)

For the odd effect error analyses, I do a slightly different analysis to what you did. For investigating errors in the control condition, I utilise a generalised linear mixed effects model - specifically, a logistic mixed effect regression. I do this for the following reasons:

1) Accuracy data is binomial data. This means it is a dichotomous dependent variable that exists in one of two states - correct or incorrect. This violates the assumptions of a standard ANOVA, thus we can model the data better if we specify the correct error distribution. 

2) When you have a dichotomous dependent variable, it is desirable to have a model and that predicts whether a value is 0 or 1. A linear regression (essentially a standard ANOVA) does not allow you to do this because the predicted y variable may not be restricted to a value between 0 and 1. Conversely, a logistic model can produce a probability score or the probability of the of the occurence of the event. 

3) The experimental design was repeated measures, which allows for the inclusion of participants as a random effect in a multi-level model. This means we can model an between participant variability in RT and thus capture more unexplained variance. Again, this will provide more statistical power. 

Below, I fit the 2 distributions to the data to demonstrate why a logistic regression is more appropriate than an ANOVA

```{r}
dist <- workingdata %>%
  dplyr::filter(condition == "controltrue")

fit_l  <- fitdist(dist$accuracy, "logis")
fit_n <- fitdist(dist$accuracy, "norm")

plot.legend <- c("logistic", "normal")
denscomp(list(fit_l, fit_n), legendtext = plot.legend)

fit_l$aic
fit_n$aic
```

To demonstrate this, I fit the accuracy data to both a logistic and normal distribution. As you can see via AIC value (lower number demonstrate better fits), the logistic distribution is a much better fit. Because of this, we should us a logistic regression rather than an ANOVA.

##  Control condition errors (lawyer only) - Odd effect error analysis (replication of slide 5)

To compute a logistic regression, I need to specify a binomial distribution. A binomial distribution is a discrete probability distribution. This means it focuses on the number of succusses within a sample which is what we need when focusing on accuracy. I also include a *logit* link function which is a transformation parameter - it is the logarithm of odds. This means we can investigate how manipulaing our indepenent variable affects the odds of scoring an accurate response. 

I use mixed effect in order to account for the repeated measures design. I specify that each participant has their own intercept (a random effect) to model for each subject.

```{r}
"Odd effect error analysis"
ggplot(workingdata %>%
         dplyr::filter(condition == "controltrue") %>%
         drop_na() %>%
         dplyr::group_by(parity) %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::summarise(errors = n()) %>%
         summarySE(measurevar = "errors", groupvars = c("parity")), aes(x = parity, y = errors)) + 
  geom_col() +
  geom_errorbar(aes(ymin = errors - ci, ymax = errors + ci), width = .2) +
  ylab("Number of errors") +
  scale_x_discrete("Parity", labels = c("odd" = "Odd", "even" = "Even")) +
  ggtitle("Odd effect - error analysis")

oddeffect_model <- glmer(accuracy ~ parity + (1 | participant),
             data = workingdata %>%
               dplyr::filter(condition == "controltrue") %>%
               drop_na(),
             family = binomial(link = "logit"))

summary(oddeffect_model)

exp(fixef(oddeffect_model))
exp(confint(oddeffect_model))

```

**Coefficients versus odds ratio**
When performing logistic regressions, coefficients can be hard to interpret due to the dichotamous nature of the data. If we exponentiate the coefficient, we get the odds-ratio.

- If an odds ratio is 1, both outcomes have an equal chance of occuring. For example if the odd ratio was 1, accurate and error responses would be equally as likely when responding with the left and right hand. 

- If an odds ratio is below 1, then a particular level of parity would decrease the chance of an accurate result. For example, an odds ratio of 0.5 would mean a level of parity has odds of 1:2, or a accurate result for every two errors.

- If an odds-ratio is larger than 1, a level of parity would increase the chance of an accurate result. For example, an odds ratio of 3 would mean a level of parity has odds of 3:1 or 3 accurate results for every 1 error.

**Interpretation of our results**

*Fixed effects*

The regression coefficient explains the change in log(odds) of the response variable (accuracy) for one unit change in the predictor variable (odd versus even numbers). In our output, odd numbers result in a decrease in the log(odds) of an accurate response. The z statistic is the equivalent of the t statistic and along with the p value, indicates that this decrease in probability of an accurate response is significant.

*Odds ratio*

As mentioned previous, this coefficient is hard to interpret. Thus we exponentiate it. This reveals that odd numbers result in an accuruate response at odds of 0.59:1, or just 59% of the time. This is significantly worse than for even numbers and demonstrates the odd effect. I have also computed confidence intervals for the odds ratio.

*Random effects*

The standard deviation of the random effects is quite high in relation to the the estimate. This suggests some substantial variability between partipants in reaction time. 

##  Control condition errors (lawyer only) - MARC effect error analysis (replication of slide 5)

```{r}
"MARC effect error analysis"
workingdata$response_hand = factor(workingdata$response_hand, levels = c("left", "right"))
workingdata$parity = factor(workingdata$parity, levels = c("odd", "even"))

ggplot(workingdata %>%
         dplyr::filter(condition == "controltrue") %>%
         drop_na() %>%
         dplyr::group_by(parity, response_hand) %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::summarise(errors = n()) %>%
         dplyr::ungroup() %>%
         summarySE(measurevar = "errors", groupvars = c("parity", "response_hand")), aes(x = parity, color = response_hand, group = response_hand, y = errors)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  geom_errorbar(aes(ymin = errors - ci, ymax = errors + ci), width = .1) +
  ylim(0, 70) +
  ylab("Number of errors") +
  xlab("Parity") +
  ggtitle("MARC effect - error analysis")

MARCeffect_model <- glmer(accuracy ~ response_hand * parity + (1 | participant),
             data = workingdata %>%
               drop_na() %>%
               dplyr::filter(condition == "controltrue") %>%
               drop_na(),
             family = binomial(link = "logit"))

summary(MARCeffect_model)

exp(fixef(MARCeffect_model))
exp(confint(MARCeffect_model))

```

**Error analysis of the MARC effect**

Because we have a significant interaction, this is what our focus should be on. As mentioned for the odd effect, the exponentiated coefficient of the main effect gives your the odds ratio.

The odd ratio for the *response_handright:parityodd* object suggests that when the response hand is right and the number is odd, the odds of an accurate response are 2.14:1, hence we would expect 2 accurate responses for every error. Intuitively this seems to be the opposite of the traditional MARC effect - we would expect responding to odd numbers with the right had to decrease the odds of an accurate response. It appears their is an inversed MARC effect. 

The line plot visualises this interaction. There are fewer errors (more accurate responses) when responding to odd numbers with the right hand. However the plot also reveals something else - this interacton appears to be driven by responding with the left hand. The number of errors with the right hand is consistent whether repsonding to even or odd numbers. However difference for the left hand is much bigger.

##  Control condition errors (lawyer only) - SNARC effect error analysis (replication of slide 5)

```{r}
# SNARC effect for errors/accuracy
workingdata$response_hand = factor(workingdata$response_hand, levels = c("right", "left"))
workingdata$magnitude = factor(workingdata$magnitude, levels = c("small", "large"))

ggplot(workingdata %>%
         dplyr::filter(condition == "controltrue") %>%
         drop_na() %>%
         dplyr::group_by(magnitude, response_hand) %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::summarise(errors = n()) %>%
         dplyr::ungroup() %>%
         summarySE(measurevar = "errors", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = errors)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  geom_errorbar(aes(ymin = errors - ci, ymax = errors + ci), width = .1) +
  ylab("Number of errors") +
  xlab("Magnitude") +
  ggtitle("SNARC effect - error analysis")

SNARCeffect_model <- glmer(accuracy ~ response_hand * magnitude + (1 | participant),
             data = workingdata %>%
               drop_na() %>%
               dplyr::filter(condition == "controltrue") %>%
               drop_na(),
             family = binomial(link = "logit"))

summary(SNARCeffect_model)

exp(fixef(SNARCeffect_model))
exp(confint(SNARCeffect_model))
```

**Error analysis of the SNARC effect**

First we calculate the odds ratio of the *response_handleft:magnitudelarge* object. This reveals that when responding with the left hand and responding to large numbers, the odds of an accurate are 2.56:1, suggesting that there is 2 correct responses for every error. Intuitively, this seems to be the opposite of the traditional SNARC effect (higher accuracy when magnitude and response hand are congruant rather than incongruant).

I implement a line plot to visualise the significant interaction. The line plot visualises this interaction. As highlighted by the logistic regression model, accuracy is higher for left + small pairs and right + large pairs. This is the opposite of the traditional SNARC effect.

**Overview**

RTs also show evidence of SNARC and MARC effect via significant ANOVAs

Logistic regression analysis for the error effects matches the analysis on slide 5 - there is a significant odd effect (increased baseline odds of errors with an odd number) however the MARC and SNARC effects are inverted. Despite overall matching results, I believe I have used a more appropriate analyse technique given the nature of the data (dichotomous variables).

Like Valter suggested, there could be trade off - decreases in reaction time lead to decreases in accuracy forthe observed effects.

Now that the lawyer only condition is analysed (both accuracy and RTs), the focus switches to experimental condition. We start with the when the lawyer is present in the "lie" experimental condition

## Lawyer lie RTs (lawyer in experimental condition) - Odd effect and MARC effect for RTs (replication of slide 7)

```{r}
### line graph - RTs over different numbers
ggplot(data = workingdata %>%
         dplyr::filter(condition == "exptrue") %>%
         dplyr::group_by(numbers, hand) %>%
         dplyr::summarise(meanRT = mean(RTs)), aes(x = numbers, y = meanRT, col = hand)) +
  geom_point() +
  geom_line() +
  scale_x_continuous("Numbers", breaks = seq(0, 9, 1)) +
  scale_y_continuous("Mean RT (s)", breaks = seq(0, 0.82, 0.02)) +
  ggtitle("RTs for left and right hands across numbers")


lawyerlie <- workingdata %>%
  dplyr::filter(condition == "exptrue")

### Odd effect - plots with 95% CIs and corresponding t-test
lawyerlie$parity <- factor(lawyerlie$parity, levels = c("odd", "even"))

ggplot(data = summarySE(lawyerlie, measurevar = "RTs", groupvars = c("parity")), aes(x = parity, y = RTs)) + 
  geom_col() +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .2) +
  coord_cartesian(ylim = c(0.62, 0.82)) +
  ggtitle("Odd effect RT analysis")

lawyerliettest <- lawyerlie %>%
  dplyr::select(RTs, parity, participant)

lawyerliettest <- reshape(lawyerliettest, idvar = "participant", timevar = "parity", direction = "wide")

t.test(lawyerliettest$RTs.even, lawyerliettest$RTs.odd, paired = TRUE)
cohen.d(lawyerliettest$RTs.even, lawyerliettest$RTs.odd, paired = TRUE)

### MARC effect - plots 95% CIs and corresponding 2x2 ANOVA
lawyerlie$parity <- factor(lawyerlie$parity, levels = c("odd", "even"))
lawyerlie$hand <- factor(lawyerlie$hand, levels = c("left", "right"))

ggplot(data = summarySE(lawyerlie, measurevar = "RTs", groupvars = c("parity", "hand")), aes(x = parity, color = hand, group = hand, y = RTs)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  ggtitle("MARC effect RT analysis")

result_MARC <- ezANOVA(lawyerlie, dv = RTs, wid = participant, within = .(parity, hand), type = 3, detailed = TRUE, return_aov = TRUE)
result_MARC

eta_sq(result_MARC$aov, partial = TRUE)
```

**Left-right line graph**

As per the lawyer control condition, we see faster RTs when parity and hand responded with are conrguent.

**Odd effect**
 
T-test reveales significant odd effect - faster reaction times for even numbers.
 
**MARC effect**

ANOVA aslo reveals significant interaction between parity and hand responded with alongside a large effect size indicating a MARC effect. 

## Laywer lie RTs (lawyer only in experimental condition) SNARC effect for RTs - replication of slide 8

```{r}
### creating magnitude variable
small <- lawyerlie %>%
  dplyr::filter(numbers <= 4) %>%
  dplyr::mutate(magnitude = "small")

large <- lawyerlie %>%
  dplyr::filter(numbers > 4) %>%
  dplyr::mutate(magnitude = "large")

lawyerlie <- rbind(small, large)

### SNARC effect - plot with 95% CIs and 2x2 ANOVA
lawyerlie$magnitude <- factor(lawyerlie$magnitude, levels = c("small", "large"))
lawyerlie$hand <- factor(lawyerlie$hand, levels = c("left", "right"))

ggplot(data = summarySE(lawyerlie, measurevar = "RTs", groupvars = c("magnitude", "hand")), aes(x = magnitude, color = hand, group = hand, y = RTs)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  ggtitle("SNARC effect RT analysis")

result_SNARC <- ezANOVA(lawyerlie, dv = RTs, wid = participant, within = .(magnitude, hand), type = 3, detailed = TRUE, return_aov = TRUE)
result_SNARC

eta_sq(result_SNARC$aov, partial = TRUE)

### regression plot

lawyerlie$numbers <- as.numeric(lawyerlie$numbers)

ggplot(data = lawyerlie %>%
         dplyr::group_by(numbers) %>%
         dplyr::summarise(m = mean(RTs)), aes(x = numbers, y = m)) +
  geom_point() +
  geom_smooth(method = "lm") +
  scale_x_continuous(breaks = seq(1, 9, 1), limits = c(1, 9)) +
  ggtitle("Linear model of number magnitude and RTs") +
  xlab("Numbers (magnitude)") +
  ylab("mean reaction time (s)")
```


Again, I make dataframes *small* and *large* in order to compute a new variable called magnitude. This will be used (alongside hand) to analyse the SNARC effect

**2x2 factorial ANOVA**

A significant interatcion with large effect size indicating a classic SNARC in the exerimental condition where a lawyer is present condition (i.e. when responding to larger numbers with right hard, RTs are faster. When responding to small numbers with left hand, RTs are faster). 

**Regression (linear model)**

Still confused about the regression model for the SNARC effect **MIGHT LEAVE OUT AND EXPLAIN TO VALTER THAT I DO NOT UNDERSTAND**

## Laywer lie RTs (lawyer only in experimental condition) Odd effect for accuracy - replication of slide 8

```{r}
### Odd effect error analysis with logistic regression
ggplot(summarySE(lawyerlie, measurevar = "accuracy", groupvars = c("parity")), aes(x = parity, y = accuracy)) + 
  geom_col() +
  geom_errorbar(aes(ymin = accuracy - ci, ymax = accuracy + ci), width = .2) +
  coord_cartesian(ylim = c(0.80, 1)) +
  ylab("Accuracy (%)") +
  scale_x_discrete("Parity", labels = c("odd" = "Odd", "even" = "Even")) +
  ggtitle("Odd effect - error analysis")

oddeffect_model <- glmer(accuracy ~ parity + (1 | participant),
             data = lawyerlie,
             family = binomial(link = "logit"))

summary(oddeffect_model)

exp(0.2769)
```

The effect of parity here refers to responding to even numbers and their accuracy when doing so. The logistic GLM indicates a significant effect of parity via the object *parityeven*. The estimate output is positive, suggesting that the even numbers significantly increase the probability of an accurate response. Plotting errors in a bar chart visualises this - even numbers have lower number of errors made. 

If we exponentiate the logistic regression estimate, we can interepret it as an odds ratio i.e. the association between exposure (even or odd numbers) and the outcome (a correct or incorrect response). Doing this indicates that the odds of a correct response increase by a factor of 1.319034 when numbers are even i.e. the correct answer is 1.3 times more likely when the number is even versus when it is odd. This statistically highlights the odd effect. 

## Laywer lie RTs (lawyer only in experimental condition) MARC effect for accuracy - replication of slide 8

```{r}
### MARC effect for errors/accuracy
lawyerlie$hand = factor(lawyerlie$hand, levels = c("left", "right"))
lawyerlie$parity = factor(lawyerlie$parity, levels = c("odd", "even"))

ggplot(data = summarySE(lawyerlie, measurevar = "accuracy", groupvars = c("parity", "hand")), aes(x = parity, color = hand, group = hand, y = accuracy)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = accuracy - ci, ymax = accuracy + ci), width = .1) +
  ylab("Response accuracy (%)") +
  xlab("Parity") +
  ggtitle("MARC effect - error analysis")

MARCeffect_model <- glmer(accuracy ~ hand * parity + (1 | participant),
             data = lawyerlie,
             family = binomial(link = "logit"))

summary(MARCeffect_model)

exp(-0.5852 )
```

**Error analysis of the MARC effect**

We focus on the significant interaction between parity and hand via the *handright:parityeven* object. Once exponentiated, this indicates that when responding to an even number, the odds ratio for people responding with their right hand decreases by a factor of 0.5569945. This is once again the opposite of the traditional MARC effect. We would expect responding to even numbers to increase the odds ratio of an accurate response when responding with the right hand. It appears their is an inversed MARC effect.

This is visualised via the line plot with 95% CIs.

## Laywer lie RTs (lawyer only in experimental condition) SNARC effect for accuracy - replication of slide 8

```{r}
lawyerlie$hand = factor(lawyerlie$hand, levels = c("right", "left"))
lawyerlie$magnitude = factor(lawyerlie$magnitude, levels = c("small", "large"))

ggplot(data = summarySE(lawyerlie, measurevar = "accuracy", groupvars = c("magnitude", "hand")), aes(x = magnitude, color = hand, group = hand, y = accuracy)) +
         stat_summary(fun.y = mean, geom = "point") +
         stat_summary(fun.y = mean, geom = "line") +
  geom_errorbar(aes(ymin = accuracy - ci, ymax = accuracy + ci), width = .1) +
  ylab("Response accuracy (%)") +
  xlab("Magnitude") +
  ggtitle("SNARC effect - error analysis")

SNARCeffect_model <- glmer(accuracy ~ hand * magnitude + (1 | participant),
             data = lawyerlie,
             family = binomial(link = "logit"))

summary(SNARCeffect_model)

exp(0.5480)
```

The *handleft:magnitudelarge* suggests that when responding with the left hand, the odd ratio for accuracy responses when responding to large numbers increases by a factor of 1.72979. Once again, ntuitively this seems to be the opposite of the traditional SNARC effect (odds ratio should decrease for large numbers  when responding with the left hand.

The line plot visualises this interaction. Accuracy is reduced for left + small pairs and right + large pairs. This is the opposite of the traditional SNARC effect.