---
title: "SNARC_MARC_analysis"
author: "Courtney Goodridge"
date: "02/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## load packages

```{r}
# rm(list = ls())

"Data visualiation"
library(ggplot2) # for plotting

"Data manipulation/loading"
library(dplyr) # data manipulation
library(tidyr) # data manipulation
library(readxl) # reading excel files 
library(openxlsx) # loads data sets from multiple sheets
library(XLConnect)

"Inferential stats and modelling"
library(car) # inferential stats
library(MASS) # inferential stats
library(EnvStats) # inferential stats
library(Rmisc) # confidence intervals 
library(sjstats) # for calculating effect sizes for ANOVA
library(effsize) # cohen's D function
library(lme4) # for linear mixed effects modelling
library(lmtest) # testing linear regression models
library(fitdistrplus) # distribution fits
library(ez) # repeated measures ANOVA 
library(insight)
library(gridExtra)
library(pdp)
```

## Theme plot

```{r}
theme_plot <-   theme(axis.title.x = element_text(size = 15), axis.text.x = element_text(size = 12), axis.title.y = element_text(size = 15), axis.text.y = element_text(size = 12), title = element_text(size = 12), legend.title = element_text(size = 15), legend.text = element_text(size = 15), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black"))
```

## load data

```{r}
dat <- lapply(excel_sheets("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/valter/Data_Lie_Numbers_Clean.xlsx"), read_excel, path = "C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/valter/Data_Lie_Numbers_Clean.xlsx")


Control_SNARC <- as.data.frame(dat[1])

Lie_SNARC <- as.data.frame(dat[2])
```

## data cleaning

```{r}
Control_SNARC <- Control_SNARC %>%
  dplyr::select(numbers, condition, key, accuracy, RTs, participant) %>%
  arrange(participant) %>%
  mutate(condition = "controltrue")

Lie_SNARC <- Lie_SNARC %>%
  dplyr::select(numbers, condition, key, accuracy, RTs, participant) %>%
  arrange(participant) %>%
  dplyr::mutate(condition = case_when(condition == "t" ~ "exptrue",
                                      condition == "l" ~ "explie"))
workingdata <- rbind(Control_SNARC, Lie_SNARC)

workingdata <- workingdata %>%
  dplyr::mutate(response_hand = case_when(key == "a" ~ "left",
                                          key == "g" ~ "right"))

workingdata <- workingdata %>%
  dplyr::mutate(parity = ifelse(numbers %% 2 == 0, "even", "odd"))

workingdata <- workingdata %>%
  dplyr::mutate(magnitude = case_when(numbers <= 4 ~ "small",
                                      numbers > 4 ~ "large"))

workingdata$magnitude <- factor(workingdata$magnitude, levels = c("small", "large"))
workingdata$response_hand <- factor(workingdata$response_hand, levels = c("left", "right"))
```

## Control condition - SNARC effect 

```{r}
controltrue <- workingdata %>%
  dplyr::filter(condition == "controltrue")

controltrue$response_hand = factor(controltrue$response_hand, levels = c("left", "right"))
controltrue$magnitude = factor(controltrue$magnitude, levels = c("small", "large"))

"plotting SNARC effect on RT"
SNARC.rt <- ggplot(controltrue %>%
         tidyr::drop_na() %>%
         summarySE(measurevar = "RTs", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = RTs)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  scale_colour_manual(name = "Response hand", labels = c("Left", "Right"), values = c("red", "blue")) + 
  ylab("RT (s)") +
  xlab("Number magnitude") +
  ggtitle("A") +
  theme_plot +
  theme(legend.position = c(0.40, 0.90), legend.key.height = unit(0, 'cm'), legend.key.width = unit(0.5, 'cm'), legend.key.size = unit(0, 'cm'), strip.text.x = element_blank(), plot.title = element_text(size = 20, face = "bold"), panel.spacing = unit(2, "lines"), axis.title.x = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8), title = element_text(size = 7), legend.title = element_text(size = 8), legend.text = element_text(size = 8))

"plotting SNARC effect on errors"
SNARC.error <- ggplot(controltrue %>%
         tidyr::drop_na() %>%
         dplyr::group_by(magnitude, response_hand) %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::summarise(errors = n()) %>%
         dplyr::ungroup() %>%
         summarySE(measurevar = "errors", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = errors)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  scale_colour_manual(name = "Response hand", labels = c("Left", "Right"), values = c("red", "blue")) +
  ylab("Number of errors") +
  xlab("Number magnitude") +
  ggtitle("B") +
  theme_plot +
  theme(legend.position = "none", legend.key.height = unit(0, 'cm'), legend.key.width = unit(0.5, 'cm'), legend.key.size = unit(0, 'cm'), strip.text.x = element_blank(), plot.title = element_text(size = 20, face = "bold"), panel.spacing = unit(2, "lines"), axis.title.x = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8), title = element_text(size = 7), legend.title = element_text(size = 8), legend.text = element_text(size = 8))

grid.arrange(SNARC.rt, SNARC.error,  nrow = 1)
setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/valter/plots")
g <- arrangeGrob(SNARC.rt, SNARC.error, nrow = 1)
  ggsave('SNARC_control.png', g,  width = 14, height = 7, units = 'cm', dpi = 300, type = 'cairo')
  
"mixed effects models"

rt.control.mod <- glmer(RTs ~ response_hand * magnitude + (1 | participant),
                        family = Gamma(link = "identity"),
                        data = controltrue)

summary(rt.control.mod)
confint(rt.control.mod)

errors.control.mod <- glmer(accuracy ~ response_hand * magnitude + (1 | participant),
                            family = binomial(link = "logit"),
                            data = controltrue %>%
                              drop_na())

summary(errors.control.mod)
```

Above I calculate plot to highlight the speed accuracy trade off between RT and errors. This is for the condition where we are just investigating the general SNARC i.e. participant is not influenced by cop or lawyer here. We see general SNARC effect for RTs but an opposite SNARC for errors, suggesting their is a speed accuracy trade-off within these data.

I also compute IES values for the data and investigate correlations between errors and RTs. Because there is an actual speed-accuracy trade off and my data is strongly correlated, the IES measure is not one that should be used. 

## Experimental condition (lie condition) - SNARC effect

```{r}
explie <- workingdata %>%
  dplyr::filter(condition == "explie")

explie$response_hand = factor(explie$response_hand, levels = c("left", "right"))
explie$magnitude = factor(explie$magnitude, levels = c("small", "large"))

"plotting SNARC effect on RT"
SNARC.rt.police <- ggplot(explie %>%
         tidyr::drop_na() %>%
         summarySE(measurevar = "RTs", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = RTs)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  scale_colour_manual(name = "Response hand", labels = c("Left", "Right"), values = c("red", "blue")) + 
  ylab("RT (s)") +
  xlab("Number magnitude") +
  ggtitle("A") +
  theme_plot +
  theme(legend.position = c(0.45, 0.95), legend.key.height = unit(0, 'cm'), legend.key.width = unit(0.5, 'cm'), legend.key.size = unit(0, 'cm'), strip.text.x = element_blank(), plot.title = element_text(size = 20, face = "bold"), panel.spacing = unit(2, "lines"), axis.title.x = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8), title = element_text(size = 7), legend.title = element_text(size = 8), legend.text = element_text(size = 8)) 

"plotting SNARC effect on errors"
SNARC.error.police <- ggplot(explie %>%
         tidyr::drop_na() %>%
         dplyr::group_by(magnitude, response_hand) %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::summarise(errors = n()) %>%
         dplyr::ungroup() %>%
         summarySE(measurevar = "errors", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = errors)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  scale_colour_manual(name = "Response hand", labels = c("Left", "Right"), values = c("red", "blue")) +
  ylab("Number of errors") +
  xlab("Number magnitude") +
  ggtitle("B") +
  theme_plot +
  theme(legend.position = "none", legend.key.height = unit(0, 'cm'), legend.key.width = unit(0.5, 'cm'), legend.key.size = unit(0, 'cm'), strip.text.x = element_blank(), plot.title = element_text(size = 20, face = "bold"), panel.spacing = unit(2, "lines"), axis.title.x = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8), title = element_text(size = 7), legend.title = element_text(size = 8), legend.text = element_text(size = 8))

setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/valter/plots")
g <- arrangeGrob(SNARC.rt.police, SNARC.error.police, nrow = 1)
  ggsave('SNARC_police.png', g,  width = 14, height = 7, units = 'cm', dpi = 300, type = 'cairo')
  
"mixed effects models"
rt.police.mod <- glmer(RTs ~ response_hand * magnitude + (1 | participant),
                        family = Gamma(link = "identity"),
                        data = explie %>%
                         drop_na())

summary(rt.police.mod)

errors.police.mod <- glmer(accuracy ~ response_hand * magnitude + (1 | participant),
                            family = binomial(link = "logit"),
                            data = explie %>%
                              drop_na())

summary(errors.police.mod)
```

In this condition, participant is shown a picture of a police officer. This means they have to lie i.e. respond with the opposite key. For example, if they shown a large number in this condition, they have to respond with their left hand. The results from the control condition are mirrored here with a standard SNARC effect for RTs but an inverse-SNARC for errors. This suggests a speed accuracy trade-off. 

## Experimental condition (truth condition) - SNARC effect

```{r}
exptrue <- workingdata %>%
  dplyr::filter(condition == "exptrue")

"plotting SNARC effect on RT"
SNARC.rt.lawyer <- ggplot(exptrue %>%
         tidyr::drop_na() %>%
         summarySE(measurevar = "RTs", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = RTs)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  geom_errorbar(aes(ymin = RTs - ci, ymax = RTs + ci), width = .1) +
  scale_colour_manual(name = "Response hand", labels = c("Left", "Right"), values = c("red", "blue")) + 
  ylab("RT (s)") +
  xlab("Number magnitude") +
  ggtitle("A") +
  theme_plot +
  theme(legend.position = c(0.50, 0.90), legend.key.height = unit(0, 'cm'), legend.key.width = unit(0.5, 'cm'), legend.key.size = unit(0, 'cm'), strip.text.x = element_blank(), plot.title = element_text(size = 20, face = "bold"), panel.spacing = unit(2, "lines"), axis.title.x = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8), title = element_text(size = 7), legend.title = element_text(size = 8), legend.text = element_text(size = 8))

"plotting SNARC effect on errors"
SNARC.error.lawyer <- ggplot(exptrue %>%
         tidyr::drop_na() %>%
         dplyr::group_by(magnitude, response_hand) %>%
         dplyr::filter(accuracy == 0) %>%
         dplyr::summarise(errors = n()) %>%
         dplyr::ungroup() %>%
         summarySE(measurevar = "errors", groupvars = c("magnitude", "response_hand")), aes(x = magnitude, color = response_hand, group = response_hand, y = errors)) +
         stat_summary(fun = mean, geom = "point") +
         stat_summary(fun = mean, geom = "line") +
  scale_colour_manual(name = "Response hand", labels = c("Left", "Right"), values = c("red", "blue")) +
  ylab("Number of errors") +
  xlab("Number magnitude") +
  ggtitle("B") +
  theme_plot +
  theme(legend.position = "none", legend.key.height = unit(0, 'cm'), legend.key.width = unit(0.5, 'cm'), legend.key.size = unit(0, 'cm'), strip.text.x = element_blank(), plot.title = element_text(size = 20, face = "bold"), panel.spacing = unit(2, "lines"), axis.title.x = element_text(size = 8), axis.text.x = element_text(size = 8), axis.title.y = element_text(size = 8), axis.text.y = element_text(size = 8), title = element_text(size = 7), legend.title = element_text(size = 8), legend.text = element_text(size = 8))

setwd("C:/Users/pscmgo/OneDrive for Business/PhD/Project/Experiment_Code/valter/plots")
g <- arrangeGrob(SNARC.rt.lawyer, SNARC.error.lawyer, nrow = 1)
  ggsave('SNARC_lawyer.png', g,  width = 14, height = 7, units = 'cm', dpi = 300, type = 'cairo')
  
"mixed effects models"
rt.lawyer.mod <- glmer(RTs ~ response_hand * magnitude + (1 | participant),
                        family = Gamma(link = "identity"),
                        data = exptrue %>%
                         drop_na())

summary(rt.lawyer.mod)

errors.lawyer.mod <- glmer(accuracy ~ response_hand * magnitude + (1 | participant),
                            family = binomial(link = "logit"),
                            data = exptrue %>%
                              drop_na())

summary(errors.lawyer.mod)
```
